---
title: "Beach Vollyball - How long will this go on?"
author: "David Scroggs"
date: "22/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(broom)
library(Hmisc)
library(visdat)

# vb_matches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-19/vb_matches.csv', guess_max = 76000) %>%
#     filter(!is.na(duration)) %>%
#   mutate(
#     Hrs = as.integer(str_extract(duration, regex("\\d+(?=\\:)"))),
#     Mins = as.integer(str_extract(duration, regex("(?<=\\:)\\d+"))),
#     Game_Time = Hrs * 60 + Mins)

#saveRDS(vb_matches,"vb_matches.RDS")

vb_matches <- readRDS("vb_matches.RDS")

theme_set(theme_minimal() + theme(plot.title = element_text(hjust = 0.5)))

```


## Introduction

I decided to start attempting a tidytuesday every other week. This is my second attempt, and first in over 2 years. I decided to blow the cobwebs off some basic stats skills that I haven't thought about in a little while and try and predict the duration of the beach volleyball games in the dataset.

DISCLAIMER: This was jus supposed to be a quick (which wasn't so quick) little forey into some old stats, and I'm sure the final conclusions don't exactly hold water. As they say, "All models are wrong ect. ect. But I'm happy for any criticism/comments.


## EDA

Quick, dirty EDA to get any idea of the types of factors that affect the game duraton. I tried to avoid including variables that would be highly correlated with the game duration (like the final score, or certain game stats). In future I'll make these decisions more formally/evidentially, but for now I just picked a few variables that piqued my interest.

### Game Durations

Game durations were parsed from the HH:MM:SS format to simply read minutes.

Game durations have a right hand tail, while the mean (red line) and median (blue line) are relatively simiar. There are some unusually long games (2 hrs+) that are outliers that stick out like the proverbials in the model checking step.

```{r}
## Game Durations
mean_dur <- mean(vb_matches$Game_Time)
median_dur <- median(vb_matches$Game_Time)

vb_matches %>% 
  ggplot(aes(x = Game_Time)) +
  geom_histogram(binwidth = 1,fill = "grey") +
  geom_vline(aes(xintercept = mean(Game_Time)), col = "red") +
  geom_vline(aes(xintercept = median(Game_Time)), col = "blue") +
  annotate(geom = "segment", x = 75, y = 2500, xend = mean_dur, yend = 2800,
           arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "label", x = 75, y = 2500, label = paste0("mean game duration is ~",round(mean_dur,0)," minutes"), hjust = "left") +
  annotate(geom = "segment", x = 75, y = 3500, xend = median_dur, yend = 3200,
           arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "label", x = 75, y = 3500, label = paste0("median game duration is ~",round(median_dur,0)," minutes"), hjust = "left") +
  scale_x_continuous(breaks = seq(0,140,15)) +
  labs(title = "Distribution of game duration",
       y = "Counts",
       x = "Game Time (minutes)")
```


There are `r sum(is.na(vb_matches$duration))` NA or undocumented game durations which are ignored for the purpose of this analysis.

### Duration by match type

There are 36 different game brackets (finals, different pool matchs, quarter finals ect.) and there is a large disparity between the number of records for the different brackets. Clearly, there will be less finals played than pool matches, but this also reflects different tournaments likely call the minor/pool rounds different names. The counts of each bracket type are shown below.

```{r}
vb_matches %>% 
  count(bracket,sort = TRUE) %>% 
  ggplot(aes(y = fct_reorder(bracket,n),x = n)) +
  geom_col(fill = "dodgerblue") +
  labs(title = "Number of matches played in each bracket",
       y = "",
       x = "Number of matches")
```

I pooled the Pool rounds to make a pooled pool... This seemed logical, as they're likely at similar points during a tournament. The updated counts in each bracket is shown below.

```{r}
vb_matches <- vb_matches %>% 
  mutate(bracket_pooled = if_else(str_detect(bracket,"Pool"),"Pool match",bracket))
  
vb_matches %>%
  count(bracket_pooled,sort = TRUE) %>% 
  ggplot(aes(y = fct_reorder(bracket_pooled,n),x = n)) +
  geom_col(fill = "dodgerblue") +
  labs(title = "Number of matches played in each bracket",
       y = "",
       x = "Number of matches")
```

Distribution of game duration for different game brackets. It appears that finals matches have a higher mean duration that other brackets.

```{r}

vb_matches %>% 
  filter(!is.na(duration)) %>% 
  group_by(bracket_pooled) %>% mutate(mean_d = mean(duration)) %>% 
  ggplot(aes(x = fct_reorder(bracket_pooled,mean_d),
             y = Game_Time,
             fill = fct_reorder(bracket_pooled,mean_d))) +
  geom_boxplot() +
  labs(title = "Distribution of game times for bracket type",
    x = "Bracket",
    y = "Game duration (mins)") +
  coord_flip()
```

Finals games are, on averge, longer than other game types. This could be explained by a greater number of sets being played in finals.

```{r}
vb_matches %>% 
  mutate(sets = str_extract_all(score,"\\d{1,2}") %>% 
              map_dbl(~length(.x)/2)) %>% 
  group_by(bracket_pooled,sets) %>% 
  summarise(n = n()) %>%
  mutate(pct = n/sum(n)) %>% 
  ggplot(aes(y = bracket_pooled, x = pct,fill = factor(sets))) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Percentage of number of sets in each bracket game",
       x = "Percentage",
       y = "",
       fill = "#Sets") +
  scale_fill_brewer(palette="Set1")
```

That doesn't appear to be the case with 3-set matches occurring at approximately same proportion as other bracket type.

### Yearly behaviour

The game duration seems relatively stable over the data set.

```{r}
vb_matches %>% 
  filter(!is.na(duration)) %>% 
  group_by(year) %>% mutate(mean_d = mean(duration)) %>% 
  ggplot(aes(x = fct_reorder(factor(year),year),
             y = Game_Time,
             fill = fct_reorder(factor(year),year))) +
  geom_boxplot() +
  #geom_vline(aes(xintercept = mean_d)) +
  labs(title = "Distribution of game times for calendar year",
    x = "Year",
    y = "Game duration (mins)") +
  coord_flip()
```

### Gender

```{r}
vb_matches %>% 
  ggplot(aes(x = Game_Time,fill = gender)) +
  geom_density(alpha = 0.5) +
   labs(title = "Distribution density of game times per gender",
    y = "Density",
    x = "Game duration (mins)")
```



```{r}
t.test(vb_matches %>% filter(gender == "W") %>% pull(Game_Time),vb_matches %>% filter(gender == "M") %>% pull(Game_Time))

```

### By Country

Where are most of the games played?

```{r}
vb_matches %>% 
  group_by(country) %>% summarise(mean = mean(Game_Time),n = n()) %>% 
  ggplot(aes(y = fct_reorder(country,mean), x = mean,
             col = n < 100, size = n)) +
  geom_vline(data = vb_matches,aes(xintercept = mean(Game_Time))) +
  geom_point()
```


### World ranking vs. duraton

```{r}
A <- vb_matches %>%
  mutate(
    w_rank = as.integer(str_extract(w_rank,"^\\d{1,3}")),
    l_rank = as.integer(str_extract(l_rank,"^\\d{1,3}")),
    d_rank = abs(w_rank - l_rank),
    ave_rank = (w_rank + l_rank)/2
    ) %>%
  filter(!is.na(w_rank),!is.na(l_rank))

A %>% 
  pivot_longer(c(w_rank,l_rank,d_rank,ave_rank),values_to = "A",names_to = "B") %>% 
  ggplot(aes(x = A, y = Game_Time)) +
  geom_point(alpha = 0.2)+
  geom_smooth(method = "lm")+
  facet_grid(B ~ .)

A.lm <- lm(data = A, Game_Time ~ w_rank + l_rank + d_rank + ave_rank)

summary(A.lm)

glance(A.lm)
```


## Age

```{r}
vb_matches %>% 
     select(matches("age")) %>% 
     reduce(`+`) %>%
     mutate(vb_matches, AGE_Sum = .,AGE_Sum = AGE_Sum/4) %>%
  filter(!is.na(AGE_Sum)) %>% select(AGE_Sum,Game_Time) %>%
  ggplot(aes(y = AGE_Sum, x = Game_Time)) +
  geom_smooth(method = "lm") +
  geom_point(alpha = 0.01)
```


## Height

```{r}
vb_matches %>% 
     select(matches("hgt")) %>% 
     reduce(`+`) %>%
     mutate(vb_matches, AGE_Sum = .,AGE_Sum = AGE_Sum/4) %>%
  filter(!is.na(AGE_Sum)) %>% select(AGE_Sum,Game_Time) %>%
  ggplot(aes(x = AGE_Sum, y = Game_Time)) +
  geom_smooth(method = "lm") +
  geom_point(alpha = 0.2)
```


## Country

```{r}

vb_matches %>% 
  group_by(country) %>% summarise(mean = mean(Game_Time),n = n()) %>% 
  ggplot(aes(y = fct_reorder(country,mean), x = mean,
             col = n < 100, size = n)) +
  geom_vline(data = vb_matches,aes(xintercept = mean(Game_Time))) +
  geom_point()

```


## Prediction

```{r}
# Helper function
tidy_summary <- function(model){
  tidy(model) %>%
  mutate(
         signif = case_when(
           p.value < 0.001 ~ "***",
           p.value < 0.01 ~ "**",
           p.value < 0.05 ~ "*",
           TRUE ~ "")) %>% 
    mutate_if(is.double,~sprintf('%.5f',.))
}
```


### Create a model tibble with altered data

```{r}
vb_model <- vb_matches %>% 
    mutate(
    ave_age = sum(w_p1_age, w_p2_age, l_p1_age, l_p2_age, na.rm = TRUE) / 4,
    ave_hgt = sum(w_p1_hgt, w_p2_hgt, l_p1_hgt, l_p2_hgt, na.rm = TRUE) / 4,
    w_rank = as.integer(str_extract(w_rank,"^\\d{1,3}")),
    l_rank = as.integer(str_extract(l_rank,"^\\d{1,3}")),
    d_rank = abs(w_rank - l_rank),
    bracket_pooled = if_else(str_detect(bracket,"Pool"),"Pool match",bracket),
    sets = str_extract_all(score,"\\d{1,2}") %>% 
              map_dbl(~length(.x)/2)
  ) %>%
  filter(!is.na(ave_age), !is.na(ave_hgt),!is.na(w_rank),!is.na(l_rank)) %>%
  mutate_at(vars(34:65),  ~ if_else(is.na(.), 0, .))
```


- Average age
- Gender
- Bracket pool
- Country
- Tournament? Lots of factors.
- Rank (Winners, loser, difference)
- Circuit
- Height
- Age


### Missing values

```{r}
vb_matches %>% select(Game_Time,contains("age"),contains("hgt"),contains("rank"),gender,bracket,country,circuit) %>%
vis_miss(warn_large_data = F)
```


```{r}

scope <- Game_Time ~ ave_age + ave_hgt + gender + bracket_pooled + country + w_rank + l_rank + d_rank + circuit

null_model <- lm(Game_Time ~ 1, data = vb_model)  

  vb.lm.m1 <- lm(null_model, data = vb_model)
vb.lm.m1 <- step(vb.lm.m1, scope = scope, direction = "both")

tidy_summary(vb.lm.m1) %>% View()
par(mfrow = c(2,2))
plot(vb.lm.m1)

```

```{r}
scope <- Game_Time ~ ave_age + ave_hgt + gender + bracket_pooled + w_rank + l_rank + d_rank + circuit

  vb.lm.m2 <- lm(null_model, data = vb_model)
vb.lm.m2 <- step(vb.lm.m2, scope = scope, direction = "both",verbose = FALSE)

summary(vb.lm.m2)
tidy_summary(vb.lm.m2) %>% View()

```

```{r}
vb_model[c(38591:38592,38584),]
```




```{r}
jonos_cv_code <- function(model, data, k = 5){
  # Split the data
data_cv <- crossv_kfold(data, k = 5)
# Fit the model to each training set
models <- map(data_cv$train, ~lm(model, data = .))
# Get predition error - gasp in awe - a function in a function.
get_pred <- function(model, test_data){
data <- as.data.frame(test_data)
pred <- add_predictions(data, model)
return(pred)
}
pred <- map2_df(models, data_cv$test, get_pred, .id = "Run")
# Get MSE
MSE <- pred %>%
group_by(Run) %>% summarise(
MSE = mean( (Game_Time - pred)^2),
n = n() )
# Get CV
CV <- sum(MSE$MSE * MSE$n) / sum(MSE$n)
return(CV)
}

glance(vb.lm.m2)$AIC

Model.CV <- tibble(
  MSE= c(
  jonos_cv_code(vb.lm.m1,vb_model),
  jonos_cv_code(vb.lm.m2,vb_model)
  ),
  AIC = c(
    glance(vb.lm.m1)$AIC,
    glance(vb.lm.m2)$AIC
    ))
Model.CV
```

