---
title: "Beach Vollyball - How long will this go on?"
author: "David Scroggs"
date: "22/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(broom)
library(Hmisc)

# vb_matches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-19/vb_matches.csv', guess_max = 76000) %>%
#     filter(!is.na(duration)) %>%
#   mutate(
#     Hrs = as.integer(str_extract(duration, regex("\\d+(?=\\:)"))),
#     Mins = as.integer(str_extract(duration, regex("(?<=\\:)\\d+"))),
#     Game_Time = Hrs * 60 + Mins)

#saveRDS(vb_matches,"vb_matches.RDS")

vb_matches <- readRDS("vb_matches.RDS")

theme_set(theme_minimal() + theme(plot.title = element_text(hjust = 0.5)))

```


## Introduction

I decided to start attempting a tidytuesday every other


## EDA

Quick, dirty EDA to get any idea of the types of factors that affect the game duraton.

### Game Durations

Game durations have a right hand tail.
```{r}
## Game Durations
mean_dur <- mean(vb_matches$Game_Time)

vb_matches %>% 
  ggplot(aes(x = Game_Time)) +
  geom_histogram(binwidth = 1,col = "black",fill = "dodgerblue") +
  geom_vline(aes(xintercept = mean(Game_Time)), col = "red") +
  annotate(geom = "curve", x = 75, y = 3000, xend = mean_dur, yend = 3000,
           arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "label", x = 75, y = 3000, label = "mean game duration is ~44 minutes", hjust = "left") +
  scale_x_continuous(breaks = seq(0,140,15)) +
  labs(title = "Distribution of game duration",
       y = "Counts",
       x = "Game Time (minutes)")
```


There are `r sum(is.na(duration))` NA or undocumented game durations.

### Duration by match type

Excluding NA durations

```{r}
vb_matches %>% 
  count(bracket,sort = TRUE) %>% 
  ggplot(aes(y = fct_reorder(bracket,n),x = n)) +
  geom_col(fill = "dodgerblue") +
  labs(title = "Number of matches played in each bracket",
       y = "",
       x = "Number of matches")
```

Theres a quite a large number of small groups that I might look at lumping together later, depending on results. These are more managable numbers. We'll see if it makes a difference when modelling.

```{r}
vb_matches <- vb_matches %>% 
  mutate(bracket_pooled = if_else(str_detect(bracket,"Pool"),"Pool match",bracket))
  
vb_matches %>%
  count(bracket_pooled,sort = TRUE) %>% 
  ggplot(aes(y = fct_reorder(bracket_pooled,n),x = n)) +
  geom_col(fill = "dodgerblue") +
  labs(title = "Number of matches played in each bracket",
       y = "",
       x = "Number of matches")
```

Distribution of game duration for different game brackets. It appears that finals matches have a higher mean duration that other brackets.

```{r}

vb_matches %>% 
  filter(!is.na(duration)) %>% 
  group_by(bracket_pooled) %>% mutate(mean_d = mean(duration)) %>% 
  ggplot(aes(x = fct_reorder(bracket_pooled,mean_d),
             y = Game_Time,
             fill = fct_reorder(bracket_pooled,mean_d))) +
  geom_boxplot() +
  labs(title = "Distribution of game times for bracket type",
    x = "Bracket",
    y = "Game duration (mins)") +
  coord_flip()
```

Finals games are, on averge, longer than other game types. This could be explained by a greater number of sets being played in finals.

```{r}
vb_matches %>% 
  mutate(sets = str_extract_all(score,"\\d{1,2}") %>% 
              map_dbl(~length(.x)/2)) %>% 
  group_by(bracket_pooled,sets) %>% 
  summarise(n = n()) %>%
  mutate(pct = n/sum(n)) %>% 
  ggplot(aes(y = bracket_pooled, x = pct,fill = factor(sets))) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Percentage of number of sets in each bracket game",
       x = "Percentage",
       y = "",
       fill = "#Sets") +
  scale_fill_brewer(palette="Set1")
```


## Yearly behaviour

```{r}
vb_matches %>% 
  filter(!is.na(duration)) %>% 
  group_by(year) %>% mutate(mean_d = mean(duration)) %>% 
  ggplot(aes(x = fct_reorder(factor(year),mean_d),
             y = Game_Time,
             fill = fct_reorder(factor(year),mean_d))) +
  geom_boxplot() +
  labs(title = "Distribution of game times for bracket type",
    x = "Bracket",
    y = "Game duration (mins)") +
  coord_flip()
```

```{r}
vb_matches %>% 
  filter(!is.na(duration)) %>% 
  ggplot(aes(x = factor(gender),
             y = Game_Time)) +
    stat_summary(fun = mean, geom = "point",position = position_dodge(width = 0.1)) + 
    stat_summary(fun.data = mean_sdl, geom = "errorbar",
                 position = position_dodge(width = 0.1),
                 fun.args = list(mult = 1), width = 0.1) +
  labs(title = "Distribution of game times per gender",
    x = "Gender",
    y = "Game duration (mins)")



```

## By Country

Where are most of the games played?

```{r}
vb_matches %>% 
  filter(!is.na(duration)) %>% 
  count(y = country) %>%
  ggplot(aes(y = fct_reorder(y,n), x = n)) +
  geom_col()


```


### World ranking vs. duraton

```{r}
vb_matches %>%
  pivot_longer(c(w_rank,l_rank),names_to = "Type",values_to = "t_ranking") %>%
  mutate(ranking = as.integer(str_extract(t_ranking,"^\\d{1,3}"))) %>%
  filter(!is.na(ranking)) %>% 
  ggplot(aes(x = as.integer(ranking),y = Game_Time)) +
  geom_point(alpha=0.2) +
  geom_smooth(method = "lm") +
  facet_grid(Type ~ .)

A <- vb_matches %>%
  mutate(
    w_rank = as.integer(str_extract(w_rank,"^\\d{1,3}")),
    l_rank = as.integer(str_extract(l_rank,"^\\d{1,3}")),
    d_rank = abs(w_rank - l_rank),
    ave_rank = (w_rank + l_rank)/2
    ) %>%
  filter(!is.na(w_rank),!is.na(l_rank))

A %>% 
  pivot_longer(c(w_rank,l_rank,d_rank,ave_rank),values_to = "A",names_to = "B") %>% 
  ggplot(aes(x = A, y = Game_Time)) +
  geom_point(alpha = 0.2)+
  geom_smooth(method = "lm")+
  facet_grid(B ~ .)

A.lm <- lm(data = A, Game_Time ~ w_rank + l_rank + d_rank + ave_rank)

summary(A.lm)

glance(A.lm)
```


## Age

```{r}
vb_matches %>% count(country)

vb_matches %>% select(contains("age")) %>% ls()

vb_matches %>% 
     select(matches("age")) %>% 
     reduce(`+`) %>%
     mutate(vb_matches, AGE_Sum = .,AGE_Sum = AGE_Sum/4) %>%
  filter(!is.na(AGE_Sum)) %>% select(AGE_Sum,Game_Time) %>%
  ggplot(aes(x = AGE_Sum, y = Game_Time)) +
  geom_smooth(method = "lm") +
  geom_point(alpha = 0.2)
```


## Height

```{r}
vb_matches %>% 
     select(matches("hgt")) %>% 
     reduce(`+`) %>%
     mutate(vb_matches, AGE_Sum = .,AGE_Sum = AGE_Sum/4) %>%
  filter(!is.na(AGE_Sum)) %>% select(AGE_Sum,Game_Time) %>%
  ggplot(aes(x = AGE_Sum, y = Game_Time)) +
  geom_smooth(method = "lm") +
  geom_point(alpha = 0.2)
```


## Country

```{r}

vb_matches %>% 
  group_by(country) %>% summarise(mean = mean(Game_Time),n = n()) %>% 
  ggplot(aes(y = fct_reorder(country,mean), x = mean,
             col = n < 100, size = n)) +
  geom_vline(data = vb_matches,aes(xintercept = mean(Game_Time))) +
  geom_point()

```


## Prediction

### Create a model tibble with altered data

```{r}
vb_model <- vb_matches %>% 
    mutate(
    ave_age = sum(w_p1_age, w_p2_age, l_p1_age, l_p2_age, na.rm = TRUE) / 4,
    ave_hgt = sum(w_p1_hgt, w_p2_hgt, l_p1_hgt, l_p2_hgt, na.rm = TRUE) / 4,
    w_rank = as.integer(str_extract(w_rank,"^\\d{1,3}")),
    l_rank = as.integer(str_extract(l_rank,"^\\d{1,3}")),
    d_rank = abs(w_rank - l_rank),
    bracket_pooled = if_else(str_detect(bracket,"Pool"),"Pool match",bracket),
    sets = str_extract_all(score,"\\d{1,2}") %>% 
              map_dbl(~length(.x)/2)
  ) %>%
  filter(!is.na(ave_age), !is.na(ave_hgt),!is.na(w_rank),!is.na(l_rank)) %>%
  mutate_at(vars(34:65),  ~ if_else(is.na(.), 0, .))
```


- Average age
- Gender
- Bracket pool
- Country
- Tournament? Lots of factors.
- Rank (Winners, loser, difference)
- Circuit
- Height
- Age


### Missing values

```{r}
vb_matches %>% select(Game_Time,contains("age"),contains("hgt"),contains("rank"),gender,bracket,country,circuit) %>%
vis_miss(warn_large_data = F)
```


```{r}

scope <- Game_Time ~ ave_age + ave_hgt + gender + bracket_pooled + country + w_rank + l_rank + d_rank + circuit

null_model <- lm(Game_Time ~ 1, data = vb_model)  

  vb.lm.m1 <- lm(null_model, data = vb_model)
vb.lm.m1 <- step(vb.lm.m1, scope = scope, direction = "both")

tidy_summary(vb.lm.m1) %>% View()

plot(vb.lm.m1)

```

```{r}
scope <- Game_Time ~ ave_age + ave_hgt + gender + bracket_pooled + w_rank + l_rank + d_rank + circuit

  vb.lm.m2 <- lm(null_model, data = vb_model)
vb.lm.m2 <- step(vb.lm.m2, scope = scope, direction = "both",verbose = FALSE)

summary(vb.lm.m2)
tidy_summary(vb.lm.m2)

```

```{r}
vb_model[c(38591:38592,38584),]
```




```{r}
jonos_cv_code <- function(model, data, k = 5){
  # Split the data
data_cv <- crossv_kfold(data, k = 5)
# Fit the model to each training set
models <- map(data_cv$train, ~lm(model, data = .))
# Get predition error - gasp in awe - a function in a function.
get_pred <- function(model, test_data){
data <- as.data.frame(test_data)
pred <- add_predictions(data, model)
return(pred)
}
pred <- map2_df(models, data_cv$test, get_pred, .id = "Run")
# Get MSE
MSE <- pred %>%
group_by(Run) %>% summarise(
MSE = mean( (Game_Time - pred)^2),
n = n() )
# Get CV
CV <- sum(MSE$MSE * MSE$n) / sum(MSE$n)
return(CV)
}

glance(vb.lm.m2)$AIC

Model.CV <- tibble(
  MSE= c(
  jonos_cv_code(vb.lm.m1,vb_model),
  jonos_cv_code(vb.lm.m2,vb_model)
  ),
  AIC = c(
    glance(vb.lm.m1)$AIC,
    glance(vb.lm.m2)$AIC
    ))
Model.CV
```


```{r}
tidy_summary <- function(model){
  tidy(model) %>%
  mutate(
         signif = case_when(
           p.value < 0.001 ~ "***",
           p.value < 0.01 ~ "**",
           p.value < 0.05 ~ "*",
           TRUE ~ ""),
  p.value = sprintf("%.5f", p.value),
         )
}

```


